# Predictive Analytics: practical 3 

## The `OJ` data set

The `OJ` data set from the `ISLR` package contains information on which of two brands of orange juice customers purchased^[The response variable is `Purchase`.] and can be loaded using


```{r}
data(OJ, package = "ISLR")
```



\noindent After loading the `caret` and `jrPred` package


```{r, message = FALSE}
library("caret")
library("jrPred")
```



\noindent make an initial examination of the relationships between each of the predictors and the response^[Use the `plot` function with a model formula or the `pairs` function.]


```{r, eval = FALSE}
par(mfrow = c(4, 5), mar = c(4, 0.5, 0.5, 0.5))
plot(Purchase ~ ., data = OJ)
```



## Initial model building using logistic regression


- To begin, create a logistic regression model that takes into consideration the prices of the two brands of orange juice, `PriceCH` and `PriceMM`. Hint: Use the `train` function, with `method = 'glm'`.  Look at the help page for the data set to understand what these variables represent.
```{r}
m1 = train(Purchase ~ PriceCH + PriceMM,
    data = OJ, method = "glm")
```
  - What proportion of purchases does this model get right?
```{r}
getTrainPerf(m1)
```
  - How does this compare to if we used no model?
```{r}
# with no model we essentially predict according to
# proportion of observations in data

# work out proportions
probs = table(OJ$Purchase)/nrow(OJ)
# sample using proportions
preds = sample(levels(OJ$Purchase), prob = probs)
# work out correct proportion
mean(preds != OJ$Purchase)
```

- Use your model to predict if a customer will buy CH or MM if the price of CH and MM is 2.3 and 2.4 respectively

```{r}
predict(m1, newdata = data.frame(PriceCH = 2.3, PriceMM = 2.4))
```

## Visualising the boundary

The `jrPred` package contains following code produces a plot of the decision boundary as seen in figure 1.

```{r, fig.cap = "Examining the decision boundary for orange juice brand purchases by price.", echo = TRUE, fig.keep="all"}
boundary_plot(m1,OJ$PriceCH, OJ$PriceMM, OJ$Purchase,
              xlab="Price CH", ylab="Price MM")
```


\noindent Run the boundary code above, and make sure you get a similar plot.

  - What happens if we add an interaction term? How does the boundary change?
```{r}
# We now have a curved decision boundary.
# There are two regions of where we would predict MM, bottom left, and a tiny one up in the top right.
```
- Try adding polynomial terms.


## Using all of the predictors


  - Instead of just using 2 predictors we want to use all of them. However, we have a few problems to tackle first. A few of our predictors are linear combinations of the others. This leads to what is called rank-deficiency problems. For instance, if you run the following model you'll realise there are a few NAs.
```{r, warning = FALSE, echo = TRUE}
mLM = train(Purchase ~ ., data = OJ, method = "glm")
```

Take the predictor PriceDiff. It is impossible to estimate it's coefficient as it is a linear combination of PriceCH and PriceMM i.e. `PriceDiff = PriceCH - PriceMM`. In this particularly data set, there are quite a few linear combinations. We can find them using the `findLinearCombos()` and `model.matrix()` functions

```{r, echo = TRUE}
remove = findLinearCombos(model.matrix(Purchase ~ ., data = OJ))
```

\noindent The output list has a component called `remove` suggesting which variables should be removed to get rid of linear combinations

```{r, echo = TRUE}
(badvar = colnames(OJ)[remove$remove])
```

We can then remove these variable from the data
```{r, echo = TRUE}
OJsub = OJ[, -remove$remove]
```


  - Use the new `OJsub` data set to model `Purchase` using all of the predictors. How accurate is the model?
```{r}
mLM = train(Purchase~., data = OJsub, method = "glm")
getTrainPerf(mLM)
```
  - What are the values of sensitivity and specificity?
```{r}
## could use confusionMatrix
(cmLM = confusionMatrix(predict(mLM,OJsub),OJsub$Purchase))
```
```{r}
# or
sensitivity(predict(mLM,OJsub),OJsub$Purchase)
specificity(predict(mLM,OJsub),OJsub$Purchase)
```
  - What does this mean?
```{r}
#The model is fairly good at picking up both positive events, person buys CH, and negative events, MM.
```

## K nearest neigbours


  - Try fitting models using the K nearest neighbours algorithm. To begin with, just have two covariates and use the `boundary_plot` function to visualise the results.

```{r, message = FALSE}
mKNN = train(Purchase~., data = OJsub, method = "knn")
```

  - How do they compare in accuracy, sensitivity and specificity?

```{r}
cmKNN = confusionMatrix(predict(mKNN,OJsub),OJsub$Purchase)
(info = data.frame(Model = c("logistic","knn"),
           Accuracy = c(cmLM$overall["Accuracy"],
               cmKNN$overall["Accuracy"]),
           Sensitivity = c(cmLM$byClass["Sensitivity"],
               cmKNN$byClass["Sensitivity"]),
           Specificity = c(cmLM$byClass["Specificity"],
               cmKNN$byClass["Specificity"])))
```

  - How does varying the number of nearest neighbours in a KNN affect the model fit?
  
```{r}
# Accuracy increases at first with knn before then getting worse after a peak value of 9.
(mKNN2 = train(Purchase~., data = OJsub, method = "knn",
    tuneGrid = data.frame(k = 1:30)))
plot(mKNN2)
```



\noindent The KNN algorithm described in the notes can also be used for regression problems. In this case the predicted response is the mean of the $k$ nearest neighbours.

  - Try fitting the KNN model for the regression problem in practical 1. You could vary the `k` parameter to try to find a better model. 
```{r, message=FALSE, warning = FALSE}
library("jrPred")
data(FuelEconomy, package = "AppliedPredictiveModeling")
regKNN = train(FE~., data = cars2010, method = "knn", tuneGrid = data.frame(k = c(1:5,10,20,50,100)))
regLM = train(FE~., data = cars2010, method = "lm")
getTrainPerf(regKNN)
getTrainPerf(regLM)
```

  - How does this compare to the linear regression models?


## An example with more than two classes

The `Glass` data set in the `mlbench` package is a data frame containing examples of the chemical analysis of $7$ different types of glass. The goal is to be able to predict which category glass falls into based on the values of the $9$ predictors.


```{r, echo = TRUE}
data(Glass, package = "mlbench")
```



\noindent A logistic regression model is typically not suitable for more than $2$ classes, so try fitting a k nearest neighbour model. Use k-fold cross validation is you want to. What proportion of predictions does your model get correct?

```{r}
tc = trainControl(method = "cv", number = 10)
model = train(Type ~ ., data = Glass, trControl = tc, method = "knn")
getTrainPerf(model)
```
